National Longitudinal Evidence for Growth in Subjective Wellbeing from Spiritual Beliefs

### 2020-11-22 Analysis.Rmd
Further within and between analysis

```{r}
table1::table1(~  Age + NZdep + Education + Employed + EthnicCategories + Male  + Partner  + Pol.Orient + Urban +  Beliefs + LIFESAT + PWI|Wave, data = dfben, overall = F)
```

@@

```{r}
dfben <- cbind(
  dfben,
  parameters::demean(dfben, select = c("Beliefs"), group = "Id")
)
```

@@

```{r models, cache = TRUE, dependson = "dfben"}
### Spirit Belief Models ### 
mod01<-lmer (LIFESAT ~ Years *(
                Beliefs__Skeptic__between  +  
                Beliefs__SpiritExcludesGod__between + 
                Beliefs_GodAndSpirit_between  + 
                Beliefs_GodExcludesSpirit_between +
                Beliefs__Skeptic__within  +  
                Beliefs__SpiritExcludesGod__within + 
                Beliefs_GodAndSpirit_within  + 
                Beliefs_GodExcludesSpirit_within) +
                (1|Id), data= dfben, na.action = na.omit)

@@

mod02<-lmer (LIFESAT ~ Years *( 
                Beliefs__Skeptic__between  +  
                Beliefs__SpiritExcludesGod__between + 
                Beliefs_GodAndSpirit_between  + 
                Beliefs_GodExcludesSpirit_between +
                Beliefs__Skeptic__within  +  
                Beliefs__SpiritExcludesGod__within + 
                Beliefs_GodAndSpirit_within  + 
                Beliefs_GodExcludesSpirit_within) +
                (1|Id), data= dfben, na.action = na.omit)

@@

# inference holds, but can we use this method with factors?  In any case, we keep to the simple model
tab_model(mod01)
tab_model(mod02)

# Inference holds 
library(lme4)
mod.1 <-lmer (PWI ~  1 +  Years * Beliefs +  Age.10yrs.C + Deprivation.S + Edu.S  + Employed + EthnicCats + Male  + Partner  + Pol.Orient.S + Urban + (1|Id), data= dfben, na.action = na.omit)
mod.2 <-lmer (LIFESAT ~ 1 + Years * Beliefs+ Age.10yrs.C + Deprivation.S + Edu.S  + Employed + EthnicCats + Male  + Partner  + Pol.Orient.S + Urban + (1|Id), data= dfben, na.action = na.omit)
summary(mod.1)
summary(mod.2)

@@

```{r model centering years}
library(lme4)
mod.1a <-lmer (PWI ~  1 +  yearsC * Beliefs +  Age.10yrs.C + Deprivation.S + Edu.S  + Employed + EthnicCats + Male  + Partner  + Pol.Orient.S + Urban + (1|Id), data= dfben)
mod.2a <-lmer (LIFESAT ~ 1 + yearsC * Beliefs+ Age.10yrs.C + Deprivation.S + Edu.S  + Employed + EthnicCats + Male  + Partner  + Pol.Orient.S + Urban + (1|Id), data= dfben)
summary(mod.1a)
summary(mod.2a)
# identical result
performance::check_collinearity(mod.1a)
performance::check_collinearity(mod.2a)
```
Centering years makes no difference to inference.

@@

>>>>

```{r}
# must remove Furniture package before running table 1
table1::table1(~  Age + NZdep + Education + Employed + EthnicCategories + Male  + Partner  + Pol.Orient + Urban +  Beliefs + LIFESAT + PWI|Wave, data = dfben, overall = F)
```

@@

```{r}
dfben <- cbind(
  dfben,
  parameters::demean(dfben, select = c("Beliefs"), group = "Id")
)

@@

```{r models, cache = TRUE, dependson = "dfben"}
### Spirit Belief Models ### 
mod01<-lmer (PWI ~ Years +
                Beliefs__Skeptic__between  +  
                Beliefs__SpiritExcludesGod__between + 
                Beliefs_GodAndSpirit_between  + 
                Beliefs_GodExcludesSpirit_between +
                Beliefs__Skeptic__within  +  
                Beliefs__SpiritExcludesGod__within + 
                Beliefs_GodAndSpirit_within  + 
                Beliefs_GodExcludesSpirit_within +
                (1|Id), data= dfben, na.action = na.omit)

parameters::check_heterogeneity(mod01, group = "ID")
parameters::check_heterogeneity(mod01, group = "ID")

mod02<-lmer (LIFESAT ~ Years +
                Beliefs__Skeptic__between  +  
                Beliefs__SpiritExcludesGod__between + 
                Beliefs_GodAndSpirit_between  + 
                Beliefs_GodExcludesSpirit_between +
                Beliefs__Skeptic__within  +  
                Beliefs__SpiritExcludesGod__within + 
                Beliefs_GodAndSpirit_within  + 
                Beliefs_GodExcludesSpirit_within +
                (1|Id), data= dfben, na.action = na.omit)

# inference holds, but can we use this method with factors? What do the contrasts mean? 
tab_model(mod01)
tab_model(mod02)
plot_models(mod01,mod02)
```

```{r models, cache = TRUE, dependson = "dfben"}
### Spirit Belief Models ### 
mod01<-lmer (PWI ~ yearsC +
                Beliefs__Skeptic__between  +  
                Beliefs__SpiritExcludesGod__between + 
                Beliefs_GodAndSpirit_between  + 
                Beliefs_GodExcludesSpirit_between +
                Beliefs__Skeptic__within  +  
                Beliefs__SpiritExcludesGod__within + 
                Beliefs_GodAndSpirit_within  + 
                Beliefs_GodExcludesSpirit_within +
                (1|Id), data= dfben, na.action = na.omit)

@@

mod02<- lmer(LIFESAT ~ yearsC +
                Beliefs__Skeptic__between  +  
                Beliefs__SpiritExcludesGod__between + 
                Beliefs_GodAndSpirit_between  + 
                Beliefs_GodExcludesSpirit_between +
                Beliefs__Skeptic__within  +  
                Beliefs__SpiritExcludesGod__within + 
                Beliefs_GodAndSpirit_within  + 
                Beliefs_GodExcludesSpirit_within +
                (1|Id), data= dfben, na.action = na.omit)

@@

# inference holds, but can we use this method with factors? What do the contrasts mean? 
tab_model(mod01)
tab_model(mod02)
plot_models(mod01,mod02)

performance::check_collinearity(mod01)
performance::check_collinearity(mod02)

# inference holds, but can we use this method with factors? What do the contrasts mean? 
tab_model(mod01)
tab_model

## BAYES

```{r bayesian_pwi_plot }
pr.LS_b <- ggpredict(model = mod.2aaB, terms = c("years [0:9]","Beliefs"), 
                    ci.lvl = 0.95,
                    type = "fe",
                    typical = "mean",
                   # condition = c(Euro =1),
                    back.transform = TRUE,
                    ppd = FALSE,
                    interval = "confidence")
p1<-plot(pr.LS_b, facets = T) + 
 # scale_y_continuous(limits=c(4.75,5.25) )  + 
  gghighlight()  +  theme_blank()+ 
  ggtitle("Bayesian Model Predicted Values of Personal Wellbeing") 
p1
pr.LS_b
```
```

```{r lme model}
# Inference holds 
library(lme4)
mod.1 <-lmer (PWI ~  1 +  Years * Beliefs +  Age.10yrs.C + Deprivation.S + Edu.S  + Employed + EthnicCats + Male  + Partner  + Pol.Orient.S + Urban + (1|Id), data= dfben, na.action = na.omit)
mod.2 <-lmer (LIFESAT ~ 1 + Years * Beliefs+ Age.10yrs.C + Deprivation.S + Edu.S  + Employed + EthnicCats + Male  + Partner  + Pol.Orient.S + Urban + (1|Id), data= dfben, na.action = na.omit)
summary(mod.1)
summary(mod.2)

@@

```{r model centering years}
library(lme4)
mod.1a <-lmer (PWI ~  1 +  yearsC * Beliefs +  Age.10yrs.C + Deprivation.S + Edu.S  + Employed + EthnicCats + Male  + Partner  + Pol.Orient.S + Urban + (1|Id), data= dfben)
mod.2a <-lmer (LIFESAT ~ 1 + yearsC * Beliefs+ Age.10yrs.C + Deprivation.S + Edu.S  + Employed + EthnicCats + Male  + Partner  + Pol.Orient.S + Urban + (1|Id), data= dfben)
summary(mod.1a)
summary(mod.2a)
# identical result
performance::check_collinearity(mod.1a)
performance::check_collinearity(mod.2a)
```

@@

```{r check collinearity}
# Note High VIFS from 
performance::check_collinearity(mod.1aa)
performance::check_collinearity(mod.2aa)

parameters::check_heterogeneity(mod.1aa, group = "ID")
parameters::check_heterogeneity(mod.2aa, group = "ID")
